Estimating the magnetization distribution within rectangular rock samples

### 2016-06-17 manuscript/manuscript.tex
final modifications for including the averaging effect produced by the finite size of the magnetic sensir

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%This modification consists in calculating, on each observation plane
%$\alpha$ (Figures \ref{fig:sample-planes} and \ref{fig:sample-planes-cross}), 
%the theoretical $\beta$-component of the magnetic induction 
%produced by each prism forming the interpretation model
%not only at the observation points
%$(x_{i}^{\alpha}, y_{i}^{\alpha}, z_{i}^{\alpha})$, $i = 1, ..., N$, 
%$\alpha = 0, 1, 2, 3$, but at the regular grid 
%of $Q$ points (open circles in Figure \ref{fig:b-mag-flux-approx}b) centered
%at the $i$th observation point $(x_{i}^{\alpha}, y_{i}^{\alpha}, z_{i}^{\alpha})$
%(black circle in Figure \ref{fig:b-mag-flux-approx}b).
%This procedure increases the number of elements of the
%vectors $\mathbf{b}_{\beta}(\mathbf{x}^{\alpha}, \mathbf{y}^{\alpha},
%\mathbf{z}^{\alpha}, \mathbf{m}^{k})$ and matrices 
%$\mathbf{M}_{\beta}^{k}(\mathbf{x}^{\alpha}, \mathbf{y}^{\alpha},
%\mathbf{z}^{\alpha})$ (equation \ref{eq:pred-data-vec-k}),
%so that they become $QN \times 1$ and $QN \times 3$,
%respectively. Similarly, the vectors
%$\mathbf{B}_{\beta}^{\alpha}(\mathbf{m})$ 
%(equation \ref{eq:pred-data-vec-alphabeta})
%become $QN \times 1$.
%Consequently, the theoretical-predicted data vector 
%$\mathbf{B}(\mathbf{m})$ (equation \ref{eq:theor-pred-data-vec-matrix})
%become $4QN \times 1$ and the matrix 
%$\mathbf{M}$ (equation \ref{eq:theor-pred-data-vec-matrix}) 
%become a $4QN \times 3P$ matrix $\tilde{\mathbf{M}}$, respectively.
%Additionally, consider that the first $Q$ elements
%of each new vectors $\tilde{\mathbf{B}}_{\beta}^{\alpha}(\mathbf{m})$
%represent the $\beta$-component of the magnetic induction
%produced by the interpretation model at the regular grid of 
%$Q$ points centered at the observation point 
%$(x_{1}^{\alpha}, y_{1}^{\alpha}, z_{1}^{\alpha})$,
%the next $Q$ elements represent the $\beta$-component
%at the regular grid of 
%$Q$ points centered at the observation point 
%$(x_{2}^{\alpha}, y_{2}^{\alpha}, z_{2}^{\alpha})$ and
%so on.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The $\beta$-component, $\beta = y, z$,
of the magnetic induction produced by the $k$-prism (in $nT$), at 
the observation points defined by the $N \times 1$ vectors 
$\mathbf{x}^{\alpha}$, $\mathbf{y}^{\alpha}$ and $\mathbf{z}^{\alpha}$,
$\alpha = 0, 1, 2, 3$,
are grouped into $N \times 1$ vectors given by
\begin{equation}
\mathbf{b}_{\beta}(\mathbf{x}^{\alpha},
                   \mathbf{y}^{\alpha},
                   \mathbf{z}^{\alpha},
                   \mathbf{m}^{k})
    = \mathbf{A}
      \mathbf{M}^{k}_{\beta}(\tilde{\mathbf{x}}^{\alpha}, 
                             \tilde{\mathbf{y}}^{\alpha},
                             \tilde{\mathbf{z}}^{\alpha}) \mathbf{m}^{k} \: ,
\label{eq:pred-data-vec-k}
\end{equation}
where $\mathbf{A}$ is a $QN \times QN$ matrix whose element $ij$ is given by
\begin{equation}
\left[\mathbf{A}\right]_{ij} = 
\begin{cases}
\begin{split}
1 &, \quad i \le j \le i+Q \\
0 &, \quad \text{otherwise}
\end{split}
\end{cases}
\label{eq:Aij}
\end{equation}
and
\begin{equation}
\mathbf{M}_{\beta}^{k}(\tilde{\mathbf{x}}^{\alpha}, 
                       \tilde{\mathbf{y}}^{\alpha},
                       \tilde{\mathbf{z}}^{\alpha}) =
\left[
\begin{array}{ccc}
\partial_{\beta x} \phi^{k}_{1} &
\partial_{\beta y} \phi^{k}_{1}&
\partial_{\beta z} \phi^{k}_{1} 
\vspace{0.05in} \\
\vdots & \vdots & \vdots \vspace{0.05in} \\
\partial_{\beta x} \phi^{k}_{N} &
\partial_{\beta y} \phi^{k}_{N} &
\partial_{\beta z} \phi^{k}_{N}
\end{array}
\right]
\label{eq:matrix-k-beta}
\end{equation}
is a $QN \times 3$ matrix whose elements $\partial_{\beta x} \phi^{k}_{j}$,
$\partial_{\beta y} \phi^{k}_{j}$ and $\partial_{\beta z} \phi^{k}_{j}$,
$j = 1, ..., QN$, 
are second derivatives, evaluated at the points defined by
the $QN \times 1$ vectors 
$\tilde{\mathbf{x}}^{\alpha}$, $\tilde{\mathbf{y}}^{\alpha}$ and 
$\tilde{\mathbf{z}}^{\alpha}$,
of the function
\begin{equation}
\phi^{k}(x, y, z) = 
    C_{m}
    \iiint\limits_{\vartheta^{k}}
    \frac{d\epsilon \, d\zeta \, d\eta}
         {\sqrt{(x - \epsilon)^{2} +
                (y - \zeta)^{2} +
                (z - \eta)^{2}}}
\label{eq:phi-k}
\end{equation}
with respect to the variables $x$, $y$ and $z$.
In equation \ref{eq:phi-k},
$C_{m}~=~10^{9}\mu_{0}/4\pi$ and $\mu_{0}$ is the 
magnetic constant and the integral is evaluated over the 
volume $\vartheta^{k}$ of the $k$th prism. 
In this work, the second derivatives in the matrices
$\mathbf{M}_{\beta}^{k}(\tilde{\mathbf{x}}^{\alpha}, 
\tilde{\mathbf{y}}^{\alpha}, \tilde{\mathbf{z}}^{\alpha})$ 
(equation \ref{eq:matrix-k-beta}) are calculated by using the formulas 
presented by \citet{nagy2000}.
Notice that the vectors $\mathbf{b}_{\beta}(\mathbf{x}^{\alpha},
\mathbf{y}^{\alpha}, \mathbf{z}^{\alpha}, \mathbf{m}^{k})$
(equation \ref{eq:pred-data-vec-k}) contain averaged values of
the $\beta$-components of the magnetic induction produced by
the interpretation model at the observation points located at
the four planes $\alpha = 0, 1, 2, 3$. The averaging is imposed
by the matrix $\mathbf{A}$ defined in equation \ref{eq:Aij}.

@@

\mathbf{A}
\mathbf{M}^{1}_{y}(\tilde{\mathbf{x}}^{3}, 
                   \tilde{\mathbf{y}}^{3},
                   \tilde{\mathbf{z}}^{3}) & 
\dots & 
\mathbf{A}
\mathbf{M}^{P}_{y}(\tilde{\mathbf{x}}^{3}, 
                   \tilde{\mathbf{y}}^{3},
                   \tilde{\mathbf{z}]^{3}) \vspace{0.05in}
\end{array}
\right] \: ,
\label{eq:sensitivity-matrix}
\end{equation}
which is formed by the matrices
$\mathbf{M}_{\beta}^{k}(\tilde{\mathbf{x}}^{\alpha}, 
\tilde{\mathbf{y}}^{\alpha}, \tilde{\mathbf{z}}^{\alpha})$, $\beta = y, z$,
$\alpha = 0,1,2,3$, $k = 1, \dots, P$ (equation \ref{eq:matrix-k-beta})
and the matrix $\mathbf{A}$ defined in equation \ref{eq:Aij}.

@@

By presuming that the rectangular rock sample can
be approximated by our previously described 
interpretation model, we define the linear inverse
problem of estimating its internal magnetization
distribution as a constrained optimization problem.
This optimization problem consists in 
estimating a specific parameter vector 
$\mathbf{m} = \mathbf{m}^{\dagger}$ (equation \ref{eq:param-vec}) 
that minimizes the goal function
\begin{equation}
\Gamma(\mathbf{m}) = \left[ 
\mathbf{d} - \mathbf{B}(\mathbf{m})
\right]^{\top}
\left[ 
\mathbf{d} - \mathbf{B}(\mathbf{m})
\right] + 
\mu_0 \, \mathbf{m}^{\top} \mathbf{R}^{\top}\mathbf{R} \mathbf{m} \: ,
\label{eq:goal-function}
\end{equation}
where the superscript``$\top$" stands for transposition, 
$\mathbf{d}$ and $\mathbf{B}(\mathbf{m})$ are,
respectively, the observed and predicted data vectors
(equations \ref{eq:obs-data-vec} and \ref{eq:theor-pred-data-vec-matrix}),
$\mu_0$ is a positive number called regularization parameter and
$\mathbf{R}$ is a sparse matrix with some elements equal to $-1$ and $1$.
For example, if the interpretation model is formed by 
two prisms ($P=2$), the matrix $\mathbf{R}$ is given by
\begin{equation}
\mathbf{R} =\begin{bmatrix}
 1& 0 & 0 & -1 & 0 & 0\\ 
0& 1 & 0 & 0 & -1 & 0\\
 0& 0 & 1 & 0 & 0 & -1\\
\end{bmatrix}_{3 \times 6}.
\label{eq:matrix_smoothness}
\end{equation} 
In a general case, $\mathbf{R}$ is a $3P-3 \times 3P$ matrix
whose $ij$th-element is given by
\begin{equation}
\left[ \mathbf{R} \right]_{ij} = 
\begin{cases}
\begin{split}
1 &\: , \quad i = j \\
-1 &\: , \quad j = i+3
\end{split}
\end{cases} \: .
\label{eq:elements_matrix}
\end{equation}
In equation \ref{eq:goal-function}, the regularization parameter $\mu_0$ 
controls the trade-off between 
the first and second terms at the right side. 
The first term represents the data fit
and the second term is the well-known first-order Tikhonov regularization 
function \citep{aster2005}, which imposes
``smoothness" to the estimated parameters.
The least-squares estimate of the parameter vector 
$\mathbf{m} = \mathbf{m}^{\dagger}$ minimizing 
the goal function (equation \ref{eq:goal-function}) is given by
\begin{equation}
\mathbf{m}^{\dagger} = \left( \mathbf{M}^{\top}\mathbf{M} + \mu_0 \,  
\mathbf{R}^T \mathbf{R}\right)^{-1}
\mathbf{M}^{\top}\mathbf{d} \: .
\label{eq:estimate}
\end{equation}
We consider that the estimated parameter vector $\mathbf{m}^{\dagger}$
given by equation \ref{eq:estimate} approximates the real magnetization 
distribution within the rectangular rock sample.

>>>>

@@

The $\beta$-component, $\beta = y, z$,
of the magnetic induction produced by the $k$-prism (in $nT$), at 
the observation points defined by the $N \times 1$ vectors 
$\mathbf{x}^{\alpha}$, $\mathbf{y}^{\alpha}$ and $\mathbf{z}^{\alpha}$,
$\alpha = 0, 1, 2, 3$,
are grouped into $N \times 1$ vectors given by
\begin{equation}
\mathbf{b}_{\beta}(\mathbf{x}^{\alpha},
                   \mathbf{y}^{\alpha},
                   \mathbf{z}^{\alpha},
                   \mathbf{m}^{k})
    = \mathbf{A}
      \mathbf{M}^{k}_{\beta}(\tilde{\mathbf{x}}^{\alpha}, 
                             \tilde{\mathbf{y}}^{\alpha},
                             \tilde{\mathbf{z}}^{\alpha}) \mathbf{m}^{k} \: ,
\label{eq:pred-data-vec-k}
\end{equation}
where $\mathbf{A}$ is a $QN \times QN$ matrix whose element $ij$,
$i = 1, ..., QN$, $j = 1, ..., QN$, is given by
\begin{equation}
\left[\mathbf{A}\right]_{ij} = 
\begin{cases}
\begin{split}
1 &, \quad (i - 1)Q \le j \le iQ \\
0 &, \quad \text{otherwise}
\end{split}
\end{cases}
\label{eq:Aij}
\end{equation}
and
\begin{equation}
\mathbf{M}_{\beta}^{k}(\tilde{\mathbf{x}}^{\alpha}, 
                       \tilde{\mathbf{y}}^{\alpha},
                       \tilde{\mathbf{z}}^{\alpha}) =
\left[
\begin{array}{ccc}
\partial_{\beta x} \phi^{k}_{1} &
\partial_{\beta y} \phi^{k}_{1}&
\partial_{\beta z} \phi^{k}_{1} 
\vspace{0.05in} \\
\vdots & \vdots & \vdots \vspace{0.05in} \\
\partial_{\beta x} \phi^{k}_{N} &
\partial_{\beta y} \phi^{k}_{N} &
\partial_{\beta z} \phi^{k}_{N}
\end{array}
\right]
\label{eq:matrix-k-beta}
\end{equation}
is a $QN \times 3$ matrix whose elements $\partial_{\beta x} \phi^{k}_{j}$,
$\partial_{\beta y} \phi^{k}_{j}$ and $\partial_{\beta z} \phi^{k}_{j}$,
$j = 1, ..., QN$, 
are second derivatives, evaluated at the points defined by
the $QN \times 1$ vectors 
$\tilde{\mathbf{x}}^{\alpha}$, $\tilde{\mathbf{y}}^{\alpha}$ and 
$\tilde{\mathbf{z}}^{\alpha}$,
of the function
\begin{equation}
\phi^{k}(x, y, z) = 
    C_{m}
    \iiint\limits_{\vartheta^{k}}
    \frac{d\epsilon \, d\zeta \, d\eta}
         {\sqrt{(x - \epsilon)^{2} +
                (y - \zeta)^{2} +
                (z - \eta)^{2}}}
\label{eq:phi-k}
\end{equation}
with respect to the variables $x$, $y$ and $z$.
In equation \ref{eq:phi-k},
$C_{m}~=~10^{9}\mu_{0}/4\pi$ and $\mu_{0}$ is the 
magnetic constant and the integral is evaluated over the 
volume $\vartheta^{k}$ of the $k$th prism. 
In this work, the second derivatives in the matrices
$\mathbf{M}_{\beta}^{k}(\tilde{\mathbf{x}}^{\alpha}, 
\tilde{\mathbf{y}}^{\alpha}, \tilde{\mathbf{z}}^{\alpha})$ 
(equation \ref{eq:matrix-k-beta}) are calculated by using the formulas 
presented by \citet{nagy2000}.
Notice that the vectors $\mathbf{b}_{\beta}(\mathbf{x}^{\alpha},
\mathbf{y}^{\alpha}, \mathbf{z}^{\alpha}, \mathbf{m}^{k})$
(equation \ref{eq:pred-data-vec-k}) contain averaged values of
the $\beta$-components of the magnetic induction produced by
the interpretation model at the observation points located at
the four planes $\alpha = 0, 1, 2, 3$. The averaging is imposed
by the matrix $\mathbf{A}$ defined in equation \ref{eq:Aij}.

@@

\mathbf{A}
\mathbf{M}^{1}_{y}(\tilde{\mathbf{x}}^{3}, 
                   \tilde{\mathbf{y}}^{3},
                   \tilde{\mathbf{z}}^{3}) & 
\dots & 
\mathbf{A}
\mathbf{M}^{P}_{y}(\tilde{\mathbf{x}}^{3}, 
                   \tilde{\mathbf{y}}^{3},
                   \tilde{\mathbf{z}}^{3}) \vspace{0.05in}
\end{array}
\right] \: ,
\label{eq:sensitivity-matrix}
\end{equation}
where $\mathbf{A}$ is defined in equation \ref{eq:Aij} and
$\mathbf{M}_{\beta}^{k}(\tilde{\mathbf{x}}^{\alpha}, 
\tilde{\mathbf{y}}^{\alpha}, \tilde{\mathbf{z}}^{\alpha})$, $\beta = y, z$,
$\alpha = 0,1,2,3$, $k = 1, \dots, P$ in equation \ref{eq:matrix-k-beta}.

@@

By presuming that the rectangular rock sample can
be approximated by our previously described 
interpretation model, we define the linear inverse
problem of estimating its internal magnetization
distribution as a constrained optimization problem.
This optimization problem consists in 
estimating a specific parameter vector 
$\mathbf{m} = \mathbf{m}^{\dagger}$ (equation \ref{eq:param-vec}) 
that minimizes the goal function
\begin{equation}
\Gamma(\mathbf{m}) = \left[ 
\mathbf{d} - \mathbf{B}(\mathbf{m})
\right]^{\top}
\left[ 
\mathbf{d} - \mathbf{B}(\mathbf{m})
\right] + 
\mu_0 \, \mathbf{m}^{\top} \mathbf{R}^{\top}\mathbf{R} \mathbf{m} \: ,
\label{eq:goal-function}
\end{equation}
where the superscript``$\top$" stands for transposition, 
$\mathbf{d}$ and $\mathbf{B}(\mathbf{m})$ are,
respectively, the observed and predicted data vectors
(equations \ref{eq:obs-data-vec} and \ref{eq:theor-pred-data-vec-matrix}),
$\mu_0$ is a positive number called regularization parameter and
$\mathbf{R}$ is a sparse matrix with some elements equal to $-1$ and $1$.
For example, if the interpretation model is formed by 
two prisms ($P=2$), the matrix $\mathbf{R}$ is given by
\begin{equation}
\mathbf{R} =\begin{bmatrix}
 1& 0 & 0 & -1 & 0 & 0\\ 
0& 1 & 0 & 0 & -1 & 0\\
 0& 0 & 1 & 0 & 0 & -1\\
\end{bmatrix}_{3 \times 6}.
\label{eq:matrix_smoothness}
\end{equation} 
In a general case, $\mathbf{R}$ is a $3P-3 \times 3P$ matrix
whose $ij$th-element is given by
\begin{equation}
\left[ \mathbf{R} \right]_{ij} = 
\begin{cases}
\begin{split}
1 &\: , \quad i = j \\
-1 &\: , \quad j = i+3 \\
0 &\: , \quad \text{otherwise}
\end{split}
\end{cases} \: .
\label{eq:elements_matrix}
\end{equation}
In equation \ref{eq:goal-function}, the regularization parameter $\mu_0$ 
controls the trade-off between 
the first and second terms at the right side. 
The first term represents the data fit
and the second term is the well-known first-order Tikhonov regularization 
function \citep{aster2005}, which imposes
``smoothness" to the estimated parameters.
The least-squares estimate of the parameter vector 
$\mathbf{m} = \mathbf{m}^{\dagger}$ minimizing 
the goal function (equation \ref{eq:goal-function}) is given by
\begin{equation}
\mathbf{m}^{\dagger} = \left( \mathbf{M}^{\top}\mathbf{M} + \mu_0 \,  
\mathbf{R}^T \mathbf{R}\right)^{-1}
\mathbf{M}^{\top}\mathbf{d} \: .
\label{eq:estimate}
\end{equation}
We consider that the estimated parameter vector $\mathbf{m}^{\dagger}$
given by equation \ref{eq:estimate} approximates the real magnetization 
distribution within the rectangular rock sample.
